Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination

### 2024-09-28 manuscript/manuscript.Rmd
Before S1 discussion

@@

modelsummary::modelsummary(
  list(`(1)` = model1, `(2)` = model2),
  vcov = vcov_list,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x Default",
    "Score" = "Performance",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
  caption = "Column 1 displays the advice given to compete against a High Performer Group based on whether the advisees were primed with low or high default and the initial preference shown to the advisor. Column 2 displays the advice given to compete against a High Performer Group based on the default and advisees' performance when their initial preference was visible to the advisor. All standard errors are clustered at the level of the advisor."
)

@@

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced
difference in expectations was too small.

>>>>

<!-- No main interactive effect on expected bonus, therefore I did not include it in this tabel -->

@@

modelsummary::modelsummary(
  list(`(1)` = model1, `(2)` = model2),
  vcov = vcov_list,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Performance",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
  caption = "Column 1 displays the advice given to compete against a High Performer Group based on whether the advisees were primed with low or high default and the initial preference shown to the advisor. Column 2 displays the advice given to compete against a High Performer Group based on the default and advisees' performance when their initial preference was visible to the advisor. All standard errors are clustered at the level of the advisor."
)

@@

Next, we examine whether advisers took the advisees' expectations into account. 
Recall that advisees who defaulted to competing against the High Performer 
Group were indeed more likely to choose this group, which, from the advisers' 
perspective, could signal high confidence in their abilities. Our theory 
predicts that showing expectations should lead to more flattering advice—
specifically, a recommendation to compete against the High Performer Group—
for advisees with a high default choice compared to those with a low 
default choice. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model with advice to compete against the high group
as the outcome measure, incorporating the default choice advisees saw, whether expectations
were shown to the adviser, and the interaction of the two. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
As predicted, we find a marginally significant interaction effect: advisees 
with a default choice to compete against the High Performer Group (nudged 
to have high expectations, though not necessarily holding them) were more 
likely to receive advice to compete against stronger competitors when 
their expectations were shown compared to when they were hidden.
We reported a linear probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee in Column 2 of Table \@ref(tab:study1regs).
We observe a significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth noting that, 
since only `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their default 
choice, this represents a conservative test of the effect of high expectations on 
the likelihood of receiving flattering advice.

```{r study1Advice fig.cap ="Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."}
study1_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#d19715", "#29526d")) +
  theme_bw() +
  labs(y = "P(Compete Against High Performer)", x = NULL) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation shown", "Expectation Absent")) +
   geom_text(aes(label=Treatment, y = -0.01), position = position_dodge(0.90))
